Determination of the \ce{^{64}Zn},\ce{^{47}Ti}(n,p) Cross-Sections using a DD Neutron Generator for Medical Isotope Studies

### 2017-02-24 First_np_Paper/first_np_paper.tex
Updated sections 2.1 and 2.2 with Lee edits

Neutron activation was carried out via irradiation in the High-Flux Neutron Generator (HFNG), a DD neutron generator at the University of California, Berkeley, seen in  \autoref{fig:hfng_a}. This generator extracts deuterium ions from an RF-heated deuterium plasma through a nozzle, whose shape was designed to form a flat-profile beam 5 mm in diameter. This deuterium beam is incident upon a water-cooled, self-loading titanium-coated copper target, where the titanium layer acts as a reaction surface for DD fusion, producing neutrons with a well-known energy distribution as a function of lab emission angle \cite{Liskien_Paulsen_1973}. Within the center of the target (at 8 mm from the DD reaction surface), samples may be inserted prior to  HFNG startup. \autoref{fig:hfng_b} displays the target setup. At the time of the presented experiments, the machine operated under nominal conditions of a 100 kV deuterium beam extracted at 1.5 mA, creating a fluence of approximately  $\sci{1.3}{7}$ neutrons/cm$^2$ on the target.

\comment{Karl: Comment or question primarily to Lee: Is it conventional to put photographs of equipment in PRC? I'm not sure it is; yes certainly for NIM, but not sure about PRC.}

\comment{Note: all images here were included at nearly full-size. They obviously need to be reduced in size, but I included them here larger the needed, so they wouldn't be pixelated if enlarged. Shrink away :)}

\begin{figure}
 \centering
%  \includegraphics[width=\columnwidth]{./figures/IMG_20160531_183154957_HDR.jpg}
 \includegraphics[scale=0.06]{./figures/IMG_20160531_183154957_HDR.jpg}
 % IMG_20160531_183154957_HDR.jpg: 2432x4320 pixel, 72dpi, 85.80x152.40 cm, bb=0 0 2432 4320
 \caption{Alternate HFNG picture - \color{red}{use this or  \autoref{fig:hfng_a}?}}
 \label{fig:alt_HFNG}
\end{figure}

\begin{figure*}
    \centering
    \begin{subfigure}[t]{0.4\textwidth}
        \centering
%         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
        \includegraphics[height=3in]{./figures/Capture.PNG}
        \caption{The UC-Berkeley High-Flux Neutron Generator,}
        \label{fig:hfng_a}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.4\textwidth}
        \centering
%         \includegraphics[width=\textwidth]{./figures/target2.png}
        \includegraphics[height=2in]{./figures/target2.png}
        \caption{Cutaway of the target chamber.}
                \label{fig:hfng_b}
    \end{subfigure}
    \caption{Note: the ion source is approximately 20 cm in diameter.}
     \label{fig:main_HFNG}
\end{figure*}

\subsection{Sample loading}\label{sec:sample_loading}

Direct activation of a thin foil can be used for high-fidelity measurement of neutron-induced reactions, through subsequent decay spectroscopy. This method can provide cross sections with far lower uncertainty than neutron time-of-flight measurements, but is heavily reliant upon precisely quantifying the neutron fluence incident upon a loaded target. To avoid the introduction of this large uncertainty, 0.5-mm thin Indium foils were co-loaded along with each target foil for irradiation, with all production cross sections thus measured relative to the well-characterized \ce{^{113}In}(n,n')\ce{^{113m}In} and \ce{^{115}In}(n,n')\ce{^{115m}In} dosimetry standard reactions. In addition, over the energy region spanned by the DD emission spectrum, the \ce{^{113}In}(n,n')\ce{^{113m}In} and \ce{^{115}In}(n,n')\ce{^{115m}In} reaction cross sections are nearly constant, which greatly simplifies the (Zn/In) and (Ti/In) ratio measurements. The use of this well-established ratio method removes the fluence dependence from all cross-section calculations, removing the largest source of systematic uncertainty in activation.

\begin{figure*}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
%         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
        \includegraphics[height=2in]{./figures/foils.png}
        \caption{Schematic (not drawn to scale) of the foil co-loading holder for the Berkeley HFNG,}
        \label{fig:holder_a}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
%         \includegraphics[width=\textwidth]{./figures/target2.png}
        \includegraphics[height=2in]{./figures/holder.pdf}
        \caption{\color{red}{Use this or \autoref{fig:holder_a} - which is more appropriate?}}
                \label{fig:holder_b}
    \end{subfigure}
     ~ 
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
%         \includegraphics[width=\textwidth]{./figures/target2.png}
        \includegraphics[height=2in]{./figures/IMG_20151103_113432563.jpg}
        \caption{ zinc (visible) foil co-loaded on top of reference indium foil, ready for irradiation.}
                \label{fig:holder_c}
    \end{subfigure}
    \caption{}
     \label{fig:main_holders}
\end{figure*}

For the measurement of the \ce{^{64}Zn}(n,p)\ce{^{64}Cu} cross section, a natural abundance commercial zinc foil of 1 cm diameter and 1 mm thickness was used as a target; a natural abundance titanium foil of the same dimensions was used as a target for the \ce{^{47}Ti}(n,p)\ce{^{47}Sc} reaction. For each run the target foils were co-loaded with a natural abundance Indium foil of 1 cm diameter and 0.5 mm thickness in a recess cut into a 2-mm thin polyethylene holder, as seen in \autoref{fig:main_holders}. Prior to loading, each foil was washed with isopropanol and dried, to remove any trace oils or residue that could become activated during irradiation.  \autoref{tab:foil_specs} records physical characteristics of each foil for the various irradiations. In each experiment, the co-loaded foils were irradiated for 3 hours at nominal operating conditions of 1.3 mA and 100 kV, to saturate the induced Indium isomer activity.

@@

The kinematics of DD fusion at 100 keV lab energy produce neutrons of energy ranging from 2.2 to 2.85 MeV, over the lab-frame emission angular range of 0-180\degree, with respect to the incident deuteron beam. This distribution is a function of the incident deuteron kinetic energy, and has been well-documented \cite{Liskien_Paulsen_1973}; \autoref{fig:scatt_angle} displays the energy-angle distribution for the incident 100 keV deuterons used in the HFNG.

\comment{Parenthetical citations, or just a list at the end? The Shimizu paper uses the latter as citation format.}

>>>>

% Neutron activation was carried out via irradiation in the High-Flux Neutron Generator (HFNG), a DD neutron generator at the University of California, Berkeley, seen in  \autoref{fig:hfng_a}. This generator extracts deuterium ions from an RF-heated deuterium plasma through a nozzle, whose shape was designed to form a flat-profile beam 5 mm in diameter. This deuterium beam is incident upon a water-cooled, self-loading titanium-coated copper target, where the titanium layer acts as a reaction surface for DD fusion, producing neutrons with a well-known energy distribution as a function of  emission angle \cite{Liskien_Paulsen_1973}. Targets are inserted in the center of the target, approximately 8 mm from the DD reaction surface, prior to startup. \autoref{fig:hfng_b} displays a cut-away schematic of the HFNG. A 100 kV deuterium beam was extracted at 1.5 mA, creating a fluence of approximately  $\sci{1.3}{7}$ neutrons/cm$^2$ on the target.

Neutron activation was carried out via irradiation in the High-Flux Neutron Generator (HFNG), a DD neutron generator at the University of California, Berkeley. This generator extracts deuterium ions from an RF-heated deuterium plasma through a nozzle, whose shape was designed to form a flat-profile beam 5 mm in diameter. This deuterium beam is incident upon a water-cooled, self-loading titanium-coated copper target, where the titanium layer acts as a reaction surface for DD fusion, producing neutrons with a well-known energy distribution as a function of  emission angle \cite{Liskien_Paulsen_1973}. Targets are inserted in the center of the target, approximately 8 mm from the DD reaction surface, prior to startup. \autoref{fig:hfng_b} displays a cut-away schematic of the HFNG. A 100 kV deuterium beam was extracted at 1.3 mA, creating a fluence of approximately  $\sci{1.3}{7}$ neutrons/cm$^2$ on the target.

% \begin{figure}
%  \centering
% %  \includegraphics[width=\columnwidth]{./figures/IMG_20160531_183154957_HDR.jpg}
%  \includegraphics[scale=0.06]{./figures/IMG_20160531_183154957_HDR.jpg}
%  % IMG_20160531_183154957_HDR.jpg: 2432x4320 pixel, 72dpi, 85.80x152.40 cm, bb=0 0 2432 4320
%  \caption{Alternate HFNG picture - \color{red}{use this or  \autoref{fig:hfng_a}?}}
%  \label{fig:alt_HFNG}
% \end{figure}

% \begin{figure*}
%     \centering
%     \begin{subfigure}[t]{0.4\textwidth}
%         \centering
% %         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
%         \includegraphics[height=3in]{./figures/Capture.PNG}
%         \caption{The UC-Berkeley High-Flux Neutron Generator,}
%         \label{fig:hfng_a}
%     \end{subfigure}%
%     ~ 
%     \begin{subfigure}[t]{0.4\textwidth}
%         \centering
% %         \includegraphics[width=\textwidth]{./figures/target2.png}
%         \includegraphics[height=2in]{./figures/target2.png}
%         \caption{Cutaway of the target chamber.}
%                 \label{fig:hfng_b}
%     \end{subfigure}
%     \caption{Note: the ion source is approximately 20 cm in diameter.}
%      \label{fig:main_HFNG}
% \end{figure*}

\begin{figure}
    \centering
        \centering
%         \includegraphics[width=\textwidth]{./figures/target2.png}
        \includegraphics[height=2in]{./figures/target2.png}
        \caption{Cut-away schematic of the HFNG. Note: the ion source is approximately 20 cm in diameter.}
                \label{fig:hfng_b}
%      \label{fig:main_HFNG}
\end{figure}

\subsection{Cross section determination by relative activation}\label{sec:sample_loading}

The approach used in both measurements was to irradiate foils of zinc or titanium, which were co-loaded with indium foils in order to determine their (n,p) cross sections relative to the well-established \ce{^{113}In}(n,n')\ce{^{113m}In} and \ce{^{115}In}(n,n')\ce{^{115m}In} neutron dosimetry standards.   \autoref{tab:foil_specs} records physical characteristics of each foil for the various irradiations. In each experiment, the co-loaded foils were irradiated for 3 hours at nominal operating conditions of 1.3 mA and 100 kV.  After irradiation, the foils were removed and placed in front of an appropriate High-Purity Germanium gamma-ray detector and the decay gamma-rays were recorded in variable-time bins.

\begin{figure}
    \centering
%         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
        \includegraphics[height=2in]{./figures/holder.pdf}
        \caption{Schematic (not drawn to scale) of the sample holder used for the Berkeley HFNG,}
        \label{fig:holder_a}
\end{figure}

% \begin{figure*}
%     \centering
%     \begin{subfigure}[t]{0.3\textwidth}
%         \centering
% %         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
%         \includegraphics[height=2in]{./figures/foils.png}
%         \caption{Schematic (not drawn to scale) of the foil co-loading holder for the Berkeley HFNG,}
%         \label{fig:holder_a}
%     \end{subfigure}%
%     ~ 
%     \begin{subfigure}[t]{0.3\textwidth}
%         \centering
% %         \includegraphics[width=\textwidth]{./figures/target2.png}
%         \includegraphics[height=2in]{./figures/holder.pdf}
%         \caption{\color{red}{Use this or \autoref{fig:holder_a} - which is more appropriate?}}
%                 \label{fig:holder_b}
%     \end{subfigure}
%      ~ 
%     \begin{subfigure}[t]{0.3\textwidth}
%         \centering
% %         \includegraphics[width=\textwidth]{./figures/target2.png}
%         \includegraphics[height=2in]{./figures/IMG_20151103_113432563.jpg}
%         \caption{ zinc (visible) foil co-loaded on top of reference indium foil, ready for irradiation.}
%                 \label{fig:holder_c}
%     \end{subfigure}
%     \caption{}
%      \label{fig:main_holders}
% \end{figure*}

1 cm diameter, 1 mm thick natural abundance zinc and titanium targets were employed for the measurement. Each of these was  co-loaded with a natural abundance Indium foil of 1 cm diameter and 0.5 mm thickness in a recess cut into a 2-mm thick polyethylene holder, as seen in \autoref{fig:holder_a}, which was mounted in the HFNG target center. Prior to loading, each foil was washed with isopropanol and dried, to remove any trace oils or residue that could become activated during irradiation.

@@

The D(D,n)\ce{^3He} reaction at 100 keV lab energy produces neutrons with energies ranging from 2.2 to 2.85 MeV, over an angular range of 0-180\degree\ in the lab frame-of-reference with respect to the incident deuteron beam. This distribution has been well documented \cite{Liskien_Paulsen_1973} and is shown in \autoref{fig:scatt_angle} for 100 keV incident deuteron energy.

\comment{These 2.2 to 2.85 MeV numbers do not quite correspond to the Figure. Addressing this with Joe.}

\comment{Is the one good reference (Liskien) sufficient?}

\comment{Should we say more about the neutron angular and energy distributions? Something from Cory's thesis or the NIM-A papers that are coming along?}

% \comment{Parenthetical citations, or just a list at the end? The Shimizu paper uses the latter as citation format.}
